---
title: "Visuals"
author: "Soumyadip"
date: "11/27/2019"
output: html_document
---
### Data Sources:
https://www.kaggle.com/new-york-state/nys-children-in-foster-care-annually
https://www.ncsc.org/Microsites/EveryKid/Home/Data-and-Reform-Efforts/Data-By-State.aspx
https://www.acf.hhs.gov/cb/resource/trends-in-foster-care-and-adoption

```{r message=FALSE, warning=FALSE}
library(readxl)
library(tidyverse)
library(viridis)
library(plotly)
library(sf)
library(leaflet)
```

## Read the data
```{r load-data-transformation}
#national dataset
nation_data<-read_excel("data/national_afcars_trends_2009_through_2018.xlsx",sheet="Data")

#State dataset
#Numbers of Children Served in Foster Care, by State
state_served <- read_excel("data/afcars_state_data_tables_09thru18.xlsx",range="Served!A8:K60") %>%
gather(year,Served,'FY 2009':'FY 2018')

#Numbers of Children in Foster Care on September 30th, by State
state_inCare <- read_excel("data/afcars_state_data_tables_09thru18.xlsx",range="In Care on September 30th!A8:K60") %>%
gather(year,InCare_Sep30,'FY 2009':'FY 2018')

#Numbers of Children Entering Foster Care, by State
state_entered <- read_excel("data/afcars_state_data_tables_09thru18.xlsx",range="Entered!A8:K60") %>%
gather(year,Entered,'FY 2009':'FY 2018')

#Numbers of Children Exiting Foster Care, by State
state_exited <- read_excel("data/afcars_state_data_tables_09thru18.xlsx",range="Exited!A8:K60") %>%
gather(year,Exited,'FY 2009':'FY 2018')

#Numbers of Children Waiting for Adoption, by State
state_waitingAdoption <- read_excel("data/afcars_state_data_tables_09thru18.xlsx",range="Waiting for Adoption!A8:K60") %>%
gather(year,Waiting_Adoption,'FY 2009':'FY 2018')

#Numbers of Children Waiting for Adoption Whose Parental Rights Have Been Terminated, by State
state_parentalRightsTerminated <- read_excel("data/afcars_state_data_tables_09thru18.xlsx",range="Parental Rights Terminated!A8:K60") %>%
gather(year,parental_rights_terminated,'FY 2009':'FY 2018')

#Numbers of Children Adopted, by State
state_adopted <- read_excel("data/afcars_state_data_tables_09thru18.xlsx",range="Adopted!A8:K60") %>%
gather(year,adopted,'FY 2009':'FY 2018')
```

## Merge the data for all categories for states
```{r merge_data}
merge_cols<-c("State","year")
#The merge argument only takes two values as input, so you have to do them separately:
#state_df<- merge(state_served,state_inCare,state_entered,state_exited,state_waitingAdoption,state_parentalRightsTerminated,state_adopted,by=c("State","year"))

state_data<- merge(state_served,state_inCare,by=merge_cols)
state_data<- merge(state_data,state_entered,by=merge_cols)
state_data<- merge(state_data,state_exited,by=merge_cols)
state_data<- merge(state_data,state_waitingAdoption,by=merge_cols)
state_data<- merge(state_data,state_parentalRightsTerminated,by=merge_cols)
state_data<- merge(state_data,state_adopted,by=merge_cols)
```

## CHeck the data
```{r}
head(state_data)
```

## Read the shape file
http://strimas.com/r/tidy-sf/
```{r}
us_states <- st_read("./shp/states.shp")
```


### Preview as a tibble
```{r}
as_tibble(us_states)
```

## Merge our data for for FY 2009
### Filter the data for 2009 and rename the state column
```{r}
state_data_2009 <- state_data %>% filter(year == 'FY 2009') %>% rename(STATE_NAME = State)
```
### Merge the data with the shape file to get the State Codes
```{r}
us_states_mapped <- inner_join(us_states,state_data_2009,by="STATE_NAME")
```


## Plot the data
https://plot.ly/r/choropleth-maps/

## check the test dataframe
```{r}
df <- read.csv("https://raw.githubusercontent.com/plotly/datasets/master/2011_us_ag_exports.csv")
head(df)
```


## Plot the data
Plotly Viridis color pallete - https://www.r-bloggers.com/how-to-use-viridis-colors-with-plotly-and-leaflet/
```{r}
#Set hover text
us_states_mapped$hover <- with(us_states_mapped,paste(STATE_NAME,'<br>',"Served: ", Served))
# give state boundaries a white border
l <- list(color = toRGB("white"), width = 2)
# specify some map projection/options
g <- list(
  scope = 'usa',
  projection = list(type = 'albers usa'),
  showlakes = TRUE,
  lakecolor = toRGB('white')
)
plot_geo(us_states_mapped, locationmode = 'USA-states') %>%
  add_trace(
    z = ~Served, text = ~hover, locations = ~STATE_ABBR,
    color = ~Served, colors = viridis_pal(option = "D")(3)
  ) %>%
  colorbar(title = "Served") %>%
  layout(
    title = 'Orphans Served by each state in 2009<br>(Hover for breakdown)',
    geo = g
  )
```
We provide viridis colors to plotly using viridis_pal and by setting option argument to “D” the default “viridis” is selected. Other options are “A” for “magma” theme, “B” – “inferno” and “C” – “plasma”. Play with letters to check which one you like the most or which suits your plot the best.


## Trying in leaflet now
https://stackoverflow.com/questions/43434898/choropleth-maps-in-r-using-leaflet-package
```{r}
popup1 <- paste0("<span style='color: #7f0000'><strong>US State Values</strong></span>",
                 "<br><span style='color: salmon;'><strong>State: </strong></span>", 
                 us_states_mapped$STATE_NAME, 
                 "<br><span style='color: salmon;'><strong>Served: </strong></span>", 
                 us_states_mapped$Served)
pal <- leaflet::colorFactor(viridis_pal(option = "D")(3), domain = us_states_mapped$Served)
leaflet(us_states_mapped) %>% 
  addProviderTiles("OpenStreetMap.Mapnik") %>%
  addPolygons(data = us_states_mapped, 
              fillColor = ~pal(Served),
              fillOpacity = 0.9,       
              color = "darkgrey",      
              weight = 1.5,            
              popup = popup1)
  # addLegend(pal = pal, values = ~Served, opacity = 0.7, title = NULL,
  # position = "bottomright")
```