--- 
title: "Foster Care Project Documentation - Visua Analytics"
author: "Soumyadip Ramya Shruti"
date: "`r Sys.Date()`"
documentclass: book
site: "bookdown::bookdown_site"
output:
  bookdown::pdf_book: default
    keep_tex: yes
link-citations: yes
bibliography: ["book.bib", "packages.bib"]
biblio-style: apalike
link-citations: yes
github-repo: soumyadipmitra/VisualAnalytics-FinalProject
description: "Detailed Project Documentation for Foster Care Shiny App Projet which is our Final Project for Visual Analytics Fall 2019"
---

# Prerequisites

This is a _sample_ book written in **Markdown**. You can use anything that Pandoc's Markdown supports, e.g., a math equation $a^2 + b^2 = c^2$.

The **bookdown** package can be installed from CRAN or Github:

```{r eval=FALSE}
install.packages("bookdown")
# or the development version
# devtools::install_github("rstudio/bookdown")
```

Remember each Rmd file contains one and only one chapter, and a chapter is defined by the first-level heading `#`.

To compile this example to PDF, you need XeLaTeX. You are recommended to install TinyTeX (which includes XeLaTeX): <https://yihui.name/tinytex/>.

```{r include=FALSE}
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown'
), 'packages.bib')
```
